---
title: "COE R Workshop"
author: "Maria Schweer-Collins, PhD & Camille Cioffi, PhD"
date: "Fall 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to R Packages 
### Installing and Loading R Packages
```{r, eval = FALSE}
# how to install R packages.
# do this the first time you use the package
install.packages("tidyverse")
# each time you open a new R session, you have to load the package
library(tidyverse)
```

### Useful R Packages
<div class="fold s o">
```{r eval = FALSE}
# data vizualization
library(tidyverse)
# APA Tables
library(apaTables)
library(sjPlot)

```
</div>


D. Introduction to workflow (XX min) 

a. R Projects 

https://r4ds.had.co.nz/workflow-projects.html 

b. Scripts, markdown, and console: differences and uses  

https://r4ds.had.co.nz/r-markdown.html 

 

**What if I already know how to use R Projects/Markdown? Review useful packages or move to Introduction to workflow 

 

E. Loading and Viewing Data (XX min) 

a. Saving your data in the same folder as your project 

b. Tibble and dataframes: uses and differences 

https://r4ds.had.co.nz/tibbles.html 

c. Tools to preview data, and summary statistics  

https://learn.datacamp.com/courses/introduction-to-the-tidyverse  

d. Hot keys: https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts 

 

G. Managing Data (For Advanced R users to move ahead) 

https://learn.datacamp.com/courses/data-manipulation-with-dplyr 

