---
title: "COE Intro to R Workshop"
author: "Maria Schweer-Collins, PhD & Camille Cioffi, PhD"
date: "Fall 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to R Packages 
### Installing and Loading R Packages
```{r, eval = FALSE}
# how to install R packages.
# do this the first time you use the package
install.packages("tidyverse")
# each time you open a new R session, you have to load the package
library(tidyverse)

# check for a package update
update.packages("tidyverse")
```

### Set Working Directory

This will tell you where your files will be saved 
```{r setwd}

# to set your working directory you can use setwd("C:/FILE PATH")

# Example
# setwd("C:/Users/jah2ax/Box Sync/_R/workshops/intro_to_R_spring_2019")



```

### Import Data into R
```{r importData}

# example od importing a .csv SPSS file

data1 <- read.csv("data.csv")


# example of importing a .sav file using package Haven
# you can also import SAS and Stata files with Haven
# install.packages("haven")
library(haven)
data2 <- read_sav("data.sav")
```

### Viewing your Data / Descriptive Statistics 

```{r}
# use package "skimr" to get the mean, standard deviation, histogram, and percentiles of numeric data
# install.packages("skimr")
library(skimr)

# use skim function to review data
skim(data1)

# use the package "psych" to get the mean, SD, range, min, max, median, skewness, and kurtosis of variables in dataset

# install.packages("psych")
library(psych)

#use describe function to review data
describe(data1)


```



### Useful R Packages
```{r eval = FALSE}
# data vizualization
library(tidyverse)
# APA Tables
library(apaTables)
library(sjPlot)

```




 


 
 

a. Saving your data in the same folder as your project 

b. Tibble and dataframes: uses and differences 

https://r4ds.had.co.nz/tibbles.html 

c. Tools to preview data, and summary statistics  

https://learn.datacamp.com/courses/introduction-to-the-tidyverse  

d. Hot keys: https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts 

 

G. Managing Data (For Advanced R users to move ahead) 

https://learn.datacamp.com/courses/data-manipulation-with-dplyr 

