---
title: "COE Intro to R Workshop"
author: "Maria Schweer-Collins, PhD & Camille Cioffi, PhD"
date: "Fall 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Introduction to R Packages 
### Installing and Loading R Packages
```{r, eval = FALSE}
# how to install R packages.
# do this the first time you use the package
install.packages("tidyverse")
# each time you open a new R session, you have to load the package
library(tidyverse)

# check for a package update
update.packages("tidyverse")

# sometimes you may receive an error that a package is not availabe for your version of R

# Installing Outdated Packages

# remotes::install_version() allows to install an outdated version of a package automatically without having to manually download the archive.

# here is an example
# install_version("MplusAutomation", version = "0.7", lib = old_lib)
```

### Set Working Directory

This will tell you where your files will be saved 
```{r setwd}
# Wondering what your working directory is set at right now?

getwd() # highlight this code and press ctrl+enter

# to set your working directory you can use setwd("C:/FILE PATH")

# Example
# setwd("C:/Users/jah2ax/Box Sync/_R/workshops/intro_to_R_spring_2019")



```

### Import Data into R
```{r importData, eval= FALSE}

# example od importing a .csv SPSS file

data1 <- read.csv("data.csv")


# example of importing a .sav file using package Haven
# you can also import SAS and Stata files with Haven
# install.packages("haven")
library(haven)
data2 <- read_sav("data.sav")

# view the top 6 rows with function head
head(data2)
```

### Viewing your Data / Descriptive Statistics 

```{r viewData, eval = FALSE}
# use package "skimr" to get the mean, standard deviation, histogram, 
# and percentiles of numeric data

# install.packages("skimr")
library(skimr)

# use skim function to review data
skim(data1)

# use the package "psych" to get the mean, SD, range, min, max, median, 
# skewness, and kurtosis of variables in dataset

# install.packages("psych")
library(psych)

#use describe function to review data
describe(data1)


```

### Ways of coding: so many ways!

```{r coding, eval = FALSE}
data2 <- data2 %>% 
  rename(weight = wt) # tidyverse way of renaming a variable name


data1$weight <- data1$wt # base R method for renaming a variable (adds a variable)
```




### Vizualizing Your Data: Basics 
```{r dataViz}
# We will use "ggplot2" to vizualize data 
#install.packackges("ggplot2)
library(ggplot2)

# The package ggplot2 comes with a dataset called mtcars already loaded

# Let's create a plot that decribes the relationship between engine size (displ) and 
# the variable fuel efficiency (hwy).

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# Do cars with bigger engines use more fuel? Look at the Plots on the left to see 
# if you think these data support that hypothesis

# here is a second example, which further divides the data by the type of vehicle (class) 

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

# Finally, you can also plot data in different facets or graphs by using something called facet_Wrap

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~class)


```




### Useful R Packages
```{r eval = FALSE}
# data vizualization
library(tidyverse)
library(ggplot2)
# APA Tables and Documents
library(apaTables)
library(sjPlot)
library(papaja) # APA manuscripts
# Others
library(shiny) # interactive web applications 
library(mice) # multiple imputation
library(lme4) # multilevel modeling
library(lavaan) # structural equation modeling 


# if you would like to read about a package, you can simply use one ? followed by the package name

# example of finding tidyverse documentation

?tidyverse

# Wondering about a function, use one ? followed by the function

?distinct

```



## Freely Available R Resources 

[Access free R tutorials here on DataCamp](https://www.datacamp.com/)

Suggested DataCamp Tutorials


* Introduction to R 
* Exploratory Data Analysis in R 
* Data Vizualization in R 


## Review the workshop flow by clicking [here](https://uvastatlab.github.io/phdplus/intror.html#get_ready) (*credit to the University of Virginia Libraries*)
 

